---
title: "cm007 Exercises: Practice with `dplyr`"
output: 
  html_document:
    keep_md: true
---

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tsibble))
```


This worksheet contains exercises aimed for practice with `dplyr`. 


_1. (a) What's the minimum life expectancy for each continent and each year? (b) Add the corresponding country to the tibble, too. (c) Arrange by min life expectancy._

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(min_life = min(lifeExp),
            country = country[lifeExp == min_life]) %>%
  arrange(min_life)
```

_2. Calculate the growth in population since the first year on record for each country by rearranging the following lines, and filling in the `FILL_THIS_IN`. Here's another convenience function for you: `dplyr::first()`._ 

```{r}
gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(rel_growth = pop - first(pop)) %>%
  DT::datatable()
```


_3. Determine the country that experienced the sharpest 5-year drop in life expectancy, in each continent, sorted by the drop, by rearranging the following lines of code. Ensure there are no `NA`'s. Instead of using `lag()`, use the convenience function provided by the `tsibble` package, `tsibble::difference()`:_

```{r}
gapminder %>% 
  group_by(country) %>%
  arrange(year) %>%
  mutate(inc_life_exp = difference(lifeExp)) %>% 
  drop_na() %>%
  ungroup() %>% 
  group_by(continent) %>%
  filter(inc_life_exp == min(inc_life_exp)) %>%
  arrange(inc_life_exp) %>%
  knitr::kable()
```

Note to self: datatable and kable don't both always knit right
